\begin{frame}{System-Level Correlation}

\begin{block}{Method}
Spearman's rank correlation (\( \rho  \)). \\
    \( \rho  \) = 1 means perfect match with humans.\\
    \( \rho  \) = -1 means perfect reverse match with humans.
\end{block}

\vfill

\begin{block}{Results}
    \begin{columns}
        \column{0.45\textwidth}
        \begin{itemize}
            \item For non-English targets: metrics relying on English tools (WordNet, POS taggers) fail. TERp and wpBleu are best there.
            \item English-German: nearly all metrics had negative correlation.
        \end{itemize}
        \column{0.45\textwidth}
\begin{table}
\centering
\begin{tabular}{l c}
\toprule
Metric & Avg rho \\
\midrule
ULC & 0.83 \\
MaxSim & 0.80 \\
RTE (absolute) & 0.79 \\
Meteor-rank & 0.75 \\
BLEU & 0.47 \\
TER / TERp & negative \\
\bottomrule
\end{tabular}
\end{table}
\end{columns}
\end{block}


\end{frame}



\begin{frame}{Sentence-Level Consistency}

\begin{block}{Method}
for every pair of systems on one sentence, did the metric agree with the human? Random baseline = 50\%.
\end{block}

\vfill

Most metrics barely beat -- or fall below -- 50\%.

\vfill

\begin{columns}

\column{0.48\textwidth}
\textbf{Into English best}

\column{0.48\textwidth}
\textbf{Out of English best}
\end{columns}
\begin{columns}
\column{0.48\textwidth}
\begin{table}
\centering
\begin{tabular}{l c}
\toprule
Metric & Score \\
\midrule
ULC & 0.54 \\
RTE & 0.53 \\
wpF & 0.52 \\
\bottomrule
\end{tabular}
\end{table}

\column{0.48\textwidth}
\begin{table}
\centering
\begin{tabular}{l c}
\toprule
Metric & Score \\
\midrule
wpF & 0.51 \\
wcd6p4er & 0.50 \\
\bottomrule
\end{tabular}
\end{table}

\end{columns}

\vfill

\begin{itemize}
    \item TER / TERp: fell BELOW 50\% -- worse than random.
\end{itemize}

\end{frame}



\begin{frame}{Sentence-Level Consistency}

\begin{block}{The ``prior'' trick}
using a metric's system-level score as a fixed prior for every sentence it produces jumps consistency to 0.57-0.61, near Oracle (0.62).
\end{block}

\vfill

\begin{block}{Insight}
which system produced the sentence predicts quality better than analyzing the sentence itself.
\end{block}

\end{frame}



\begin{frame}{Key Takeaways}

\begin{itemize}
    \item ULC best predicts human judgment at system level. BLEU is overused and underperforms.
    \item For non-English targets, TERp and wpBleu hold up; most others break down.
    \item Sentence-level evaluation is extremely hard -- most metrics are near coin-flip.
    \item Using system-level performance as a prior significantly improves sentence-level consistency -- a clear direction for future work.
\end{itemize}

\end{frame}
