\begin{frame}{Translation Task Results}
In this section, we will try to answer these three questions:

\begin{itemize}
    \item Which systems produced the best translation quality for each language pair?
    \item Did the system combinations produce better translations than individual systems?
    \item Which of the systems that used only the provided training materials produced the best translation quality?
\end{itemize}
\end{frame}

\begin{frame}{Which MT systems are best?}
\textbf{How “Best” is Defined?}
\begin{itemize}
    \item Based on human evaluation (sentence ranking)
    \item Systems ranked by:
    \item How often their translations are better than or equal to others
    \item They used Sign Test (p \( \leq \) 0.1)
\end{itemize}

\vspace{0.5em}

\begin{table}
\centering
    \begin{tabularx}{\textwidth}{lLLL}
\toprule
Pair & Performance Range & Top systems (C) & Top systems (NC) \\
\midrule
French $\rightarrow$ English & 0.76 - 0.34 & DCU, JHU & Google, LIMSI \\
English $\rightarrow$ French & 0.73 - 0.38 & UKA, DCU & LIUM-SYSTRAN, Google \\
\bottomrule
\end{tabularx}
\end{table}
\end{frame}

\begin{frame}{Which MT systems are best?}
\begin{table}
\centering
    \begin{tabularx}{\textwidth}{lLLL}
\toprule
Pair & Performance Range & Top systems (C) & Top systems (NC) \\
\midrule
German $\rightarrow$ English & 0.66 - 0.33 & STUTTGART, UEDIN & RBMT5, USAAR \\
English $\rightarrow$ German & 0.66 - 0.43 & UKA, UEDIN & RBMT2, RBMT3 \\
Czech $\rightarrow$ English & 0.75 - 0.51 & UEDIN, CU-BOJAR & Google \\
Hungarian $\rightarrow$ English & 0.75 - 0.45 & UMD, UEDIN & MORPHO \\
\bottomrule
\end{tabularx}
\end{table}
\end{frame}

\begin{frame}{Which MT systems are best?}
\begin{table}
\centering
    \begin{tabularx}{\textwidth}{lLLL}
\toprule
Pair & Performance Range & Top systems (C) & Top systems (NC) \\
\midrule
Spanish $\rightarrow$ English & 0.70 - 0.37 & TALP-UPC, WEDIN & Google, RBMT1 \\
English $\rightarrow$ Spanish & 0.66 - 0.48 & UEDIN, NUS & RBMT3, Google \\
English $\rightarrow$ Czech & 0.67 - 0.48 & CU-BOJAR, WEDIN & PCTRANS, EUROTRANXP \\
\bottomrule
\end{tabularx}
\end{table}
\end{frame}

\begin{frame}{System Combinations vs Individual Systems}
    \footnotesize
    \begin{table}[htbp]
\centering
        \begin{tabular}{l P{1.7in} c P{2in}}
\toprule
\textbf{Language Pair} & \textbf{Best system combinations} & \textbf{Entries} & \textbf{Significantly different than best individual systems?} \\
\midrule

German-English 
& RWTH-COMBO, BBN-COMBO, CMU-COMBO, USAAR-COMBO 
& 5 
& BBN-COMBO $>$ GOOGLE, SYSTRAN; USAAR-COMBO $<$ RMBT2; no difference for others \\

English-German 
& USAAR-COMBO 
& 1 
& worse than 3 best systems \\

Spanish-English 
& CMU-COMBO, USAAR-COMBO, BBN-COMBO 
& 3 
& each better than one of the RBMT systems, but there was no difference with GOOGLE, TALP-UPC \\

English-Spanish 
& USAAR-COMBO 
& 1 
& no difference \\

French-English 
& CMU-COMBO-HYPOSEL, DCU-COMBO, CMU-COMBO 
& 5 
& no difference \\

English-French 
& USAAR-COMBO, DCU-COMBO 
& 2 
& USAAR-COMBO $>$ UKA; DCU-COMBO $>$ SYSTRAN, LIMSI; no difference with others \\

Czech-English 
& CMU-COMBO 
& 2 
& no difference \\

Hungarian-English 
& CMU-COMBO-HYPOSEL, CMU-COMBO 
& 3 
& both worse than MORPHO \\

Multisource-English 
& RWTH-COMBO 
& 3
& N/A \\

\bottomrule
\end{tabular}
\end{table}
\end{frame}

\begin{frame}{Are system combinations better than single systems?}
\begin{block}{Understandability}
\begin{itemize}
    \item Measures how often translated output is clear and meaningful after editing
    \item Based on human judgment (Yes/No) on edited translations
    \item Reflects real-world usability of translations
\end{itemize}
\end{block}

\vfill
\textbf{Results:}
\begin{itemize}
    \item ~50\% acceptable translations (French - English, German - English, English - Spanish)
    \item ~40\% acceptable translations (Spanish - English)
    \item ~30\% acceptable translations (English - Czech)
    \item ~20\% acceptable translations (Czech - English, Hungarian - English)
\end{itemize}

\vfill
\begin{itemize}
    \item Even top systems are only correct ~half the time
    \item Large variation across language pairs
\end{itemize}
\end{frame}

\begin{frame}{Takeaways}
\begin{itemize}
    \item Human evaluation is the gold standard
    \item Commercial systems (Google, RBMT) perform strongly
    \item Performance drops for complex/hard language pairs
    \item Many systems perform similarly $\rightarrow$ hard to declare a clear winner
    \item System combinations doesn’t guarantee improvement
    \item Machine Translation quality is still far from perfect
\end{itemize}
\end{frame}
